---
title: Modeling Relationship Between Full Carrier Risk and Carrier Risk with Unaffected
  Relative Information Masked
output:
  html_document:
    df_print: paged
  pdf_document: default
---



```{r setup}
library(tidyverse)
library(MASS)

numberFamilies <- 10000

#load in the families dataframe
load(str_interp("../RObjects/family_dfs/listOfDataFrames/familyinfo${numberFamilies}Families_no_for_each.Rdata"))

#load in the summary table of PanelPro results
load(str_interp("../RObjects/summary_tables/summaryTable${numberFamilies}Families.Rdata"))

probandAge <- c()
for(i in 1:length(families)){
  proband = families[[i]] %>% filter(isProband==1) 
  probandAge <- c(probandAge, proband$CurAge)
}
summaryTable$probandAge <- probandAge
```


```{r remove carrier risk of 1}
summaryTable<- summaryTable %>% filter(carrierRiskUnaffectedInfoMasked < 1)
summaryTable<- summaryTable %>% filter(fullCarrierRisk < 1)
print(nrow(summaryTable))
```

```{r logit transformation}
logit <- function(x){
  log(x)-log(1-x)
}

#offset <- 0.00001
summaryTable$logitFullCarrierRisk <- logit(summaryTable$fullCarrierRisk)
summaryTable$logitCarrierRiskUnaffectedInfoMasked <- logit(summaryTable$carrierRiskUnaffectedInfoMasked)
#logit spreads out those lower values and higher values
```

```{r add indicator variables}
summaryTable$numAffectedRelativesIndicator0 <- ifelse(summaryTable$numAffectedRelatives == 0, 1, 0)
summaryTable$numAffectedRelativesIndicator1 <- ifelse(summaryTable$numAffectedRelatives == 1, 1, 0)
summaryTable$numAffectedRelativesIndicator2 <- ifelse(summaryTable$numAffectedRelatives == 2, 1, 0) 
summaryTable$numAffectedRelativesIndicator3 <- ifelse(summaryTable$numAffectedRelatives >= 3, 1, 0) 

summaryTable$group <- ifelse(summaryTable$numAffectedRelativesIndicator0 == 1, "0",
                                  ifelse(summaryTable$numAffectedRelativesIndicator1 == 1, "1",
                                  ifelse(summaryTable$numAffectedRelativesIndicator2 == 1, "2",
                                  ifelse(summaryTable$numAffectedRelativesIndicator3 == 1, "3+","NA"))))

#checking for correlations that may induce collinearity
cor_df <- summaryTable[, c("logitFullCarrierRisk", "logitCarrierRiskUnaffectedInfoMasked", "numAffectedRelativesIndicator1", "numAffectedRelativesIndicator2","numAffectedRelativesIndicator3")] 
print(cor(cor_df))

```

# Families with First Degree Affected Relatives
```{r restrict to families with first degree affected relatives}
summaryTableFirstDegAff <- summaryTable %>% filter(firstDegreeAffectedFamilyMembersBinary == 1)
nrow(summaryTableFirstDegAff)
```

## Using Indicator Variables



```{r fit linear model first degree aff with indicators}
#take out the 0 indicator because we are restricting to families with >= 1 affected relative

fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked  + numAffectedRelativesIndicator1 + numAffectedRelativesIndicator2, data=summaryTableFirstDegAff)
print(summary(fitLinear))
```





```{r plot the model first degree aff with indicators}


ggplot(data = summaryTableFirstDegAff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point() +
  geom_line(aes(y = predict(fitLinear), color = as.factor(group))) +
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression", color = "Number of Affected Relatives") 


```

```{r view residuals first degree aff with indicators}
res <- resid(fitLinear)
plot(fitted(fitLinear), res)
abline(0,0)
```
```{r indicators model evaluation first deg aff}
# get predicted values for training data
expit <- function(x){
  exp(x) / (1 + exp(x))
}

summaryTableFirstDegAff$predicted <- expit(predict(fitLinear)) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegAff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegAff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```




## Without Using Indicator Variables
```{r fit linear model first degree aff}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked , data=summaryTableFirstDegAff)
print(summary(fitLinear))
```


```{r plot the model first degree aff}
ggplot(data = summaryTableFirstDegAff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression", color = "Proband MLH1")

```


```{r view residuals first degree aff}
res <- resid(fitLinear)
plot(fitted(fitLinear), res)
abline(0,0)
```

```{r qqplot first degree aff}
# make a qq plot to assess normality

qqnorm(res)
qqline(res) 
```


```{r model evaluation first deg aff}
# get predicted values for training data
summaryTableFirstDegAff$predicted <- expit(predict(fitLinear)) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegAff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegAff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```
## Testing Covariates: Number of Affected Relatives, Age of Proband


### Covariate: Number of Affected Relatives
```{r fit linear model first degree aff covariate num aff rels}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked + numAffectedRelatives , data=summaryTableFirstDegAff)
print(summary(fitLinear))
```


```{r plot the model first degree aff covariate num aff rels}
ggplot(data = summaryTableFirstDegAff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color = as.factor(numAffectedRelatives))) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression", subtitle="First Degree Family Members Affected, Covariate: Number of Affected Relatives", color = "Number of Affected Relatives")

```

```{r model evaluation first deg aff covariate number of affected relatives}
# get predicted values for training data
summaryTableFirstDegAff$predicted <- expit(predict(fitLinear)) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegAff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegAff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```


### Covariate: Proband Age

```{r fit linear model first degree aff covariate proband age}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked + probandAge , data=summaryTableFirstDegAff)
print(summary(fitLinear))
```


```{r plot the model first degree aff covariate proband age}
ggplot(data = summaryTableFirstDegAff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color =probandAge)) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression",subtitle="First Degree Family Members Affected, Covariate: Proband Age", color = "Proband Age")

```


### Covariates: Number of Affected Relatives and Proband Age

```{r fit linear model first degree aff covariates num aff rels and proband age}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked + probandAge + numAffectedRelatives , data=summaryTableFirstDegAff)
print(summary(fitLinear))
```
Proband age again seems statistically insignificant

```{r plot the model first degree aff covariates num aff rels and proband age}
ggplot(data = summaryTableFirstDegAff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point() +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression",subtitle="First Degree Family Members Affected, Covariates: Proband Age and Number of Affected Relatives", color = "Proband Age")

```

# Families with No Affected First Degree Relatives

```{r filter to no affected first degree relatives}
summaryTableFirstDegUnaff <- summaryTable %>% filter(firstDegreeAffectedFamilyMembersBinary == 0)
nrow(summaryTableFirstDegUnaff)
```

## Using Indicator Variables

```{r fit linear model first degree unaff with indicators}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked  + numAffectedRelativesIndicator0 + numAffectedRelativesIndicator1 + numAffectedRelativesIndicator2, data=summaryTableFirstDegUnaff)
print(summary(fitLinear))
```

```{r plot the model first degree unaff with indicators}


ggplot(data = summaryTableFirstDegUnaff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point() +
  geom_line(aes(y = predict(fitLinear), color = as.factor(group))) +
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression", color = "Number of Affected Relatives") 
```

```{r view residuals first degree unaff with indicators}
res <- resid(fitLinear)
plot(fitted(fitLinear), res)
abline(0,0)
```
```{r indicators model evaluation first deg unaff}
# get predicted values for training data
expit <- function(x){
  exp(x) / (1 + exp(x))
}

summaryTableFirstDegUnaff$predicted <- expit(predict(fitLinear))

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegUnaff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegUnaff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```



## Without Using Indicator Variables



```{r fit linear model first deg unaff}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked , data=summaryTableFirstDegUnaff)
print(summary(fitLinear))
```

```{r plot the model first deg unaff}
ggplot(data = summaryTableFirstDegUnaff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color=probandMLH1Status)) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information Linear Regression", subtitle="First Degree Family Members Unaffected", color = "Proband MLH1")
```

```{r view residuals first deg unaff}
res <- resid(fitLinear)
plot(fitted(fitLinear), res)
abline(0,0)
```



```{r qqplot first degree unaff}
# make a qq plot to assess normality

qqnorm(res)
qqline(res) 
```


```{r model evaluation first deg unaff}
# get predicted values for training data

summaryTableFirstDegUnaff$predicted <- expit(predict(fitLinear)) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegUnaff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegUnaff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```
## Testing Covariates: Number of Affected Relatives, Age of Proband


### Covariate: Number of Affected Relatives
```{r fit linear model first deg unaff covariate num aff rels}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked + numAffectedRelatives , data=summaryTableFirstDegUnaff)
print(summary(fitLinear))
```


```{r plot the model first deg unaff covariate num aff rels}
ggplot(data = summaryTableFirstDegUnaff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color = as.factor(numAffectedRelatives))) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression", subtitle="First Degree Family Members Unaffected, Covariate: Number of Affected Relatives", color = "Number of Affected Relatives")

```
### Covariate: Proband Age




```{r fit linear model first deg unaff covariate proband age}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked + probandAge , data=summaryTableFirstDegUnaff)
print(summary(fitLinear))
```
Here the proband age seems more statistically signficant for families that do not have an affected first degree relative

```{r plot the model  first deg unaff covariate proband age}
ggplot(data = summaryTableFirstDegUnaff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color =probandAge)) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression",subtitle="First Degree Family Members Unaffected, Covariate: Proband Age", color = "Proband Age")

```


### Covariates: Number of Affected Relatives and Proband Age

```{r fit linear model first deg unaff covariates num aff rels and proband age}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked + probandAge + numAffectedRelatives , data=summaryTableFirstDegUnaff)
print(summary(fitLinear))
```
Proband age seems statistically significant here as well

```{r plot the model first deg unaff covariates num aff rels and proband age}
ggplot(data = summaryTableFirstDegUnaff, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point() +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Linear Regression",subtitle="First Degree Family Members Unaffected, Covariates: Proband Age and Number of Affected Relatives", color = "Proband Age")

```

```{r model evaluation first deg unaff covariates}
# get predicted values for training data

summaryTableFirstDegUnaff$predicted <- expit(predict(fitLinear)) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegUnaff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegUnaff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```

# All families

First we will run a linear regression model on all families generated. 



```{r linear model all families}
fitLinear <- lm(logitFullCarrierRisk ~ logitCarrierRiskUnaffectedInfoMasked , data=summaryTable)
print(summary(fitLinear))

```

```{r residuals of lm all families}
residuals <- residuals(fitLinear)
qqnorm(residuals)
qqline(residuals)
#the residuals should fit to the line in the qqplot if the data is normally distributed

range(summaryTable$fullCarrierRisk)
range(summaryTable$carrierRiskUnaffectedInfoMasked)
range(summaryTable$logitFullCarrierRisk)
range(summaryTable$logitCarrierRiskUnaffectedInfoMasked)
```

```{r plot all families}
ggplot(data = summaryTable, aes(x= logitCarrierRiskUnaffectedInfoMasked, y= logitFullCarrierRisk)) +
  geom_point(aes(color = as.factor(firstDegreeAffectedFamilyMembersBinary))) +
  geom_line(aes(y = predict(fitLinear)))+
  labs(x = "Carrier Risk Unaffected Info Masked (logit)", y= "Carrier Risk for Full Family Information (logit)", title = "Comparing Carrier Risk for Masked Info and Full Family Information with Linear Regression", color = "1st Deg Aff")

```

```{r model evaluation all families}
# get predicted values for training data

summaryTable$predicted <- expit(predict(fitLinear)) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTable$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTable$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTable$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```

# Other models

## Loess Regression

### All Families

```{r loess model fit and plot}
fitLoess <- loess(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked, data = summaryTable, span = 0.2)
summary(fitLoess) #RSE is 0.00397 (seems good but I'm not sure)

loessVis = ggplot(data = summaryTable, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitLoess)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "Comparing Carrier Risk for Masked Info and Full Family Information with Loess Regression")

print(loessVis)
```

```{r model evaluation loess all families}
# get predicted values for training data

summaryTable$predicted <- predict(fitLoess) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTable$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTable$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTable$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```


### First Degree Affected Families

```{r loess model fit and plot first deg aff}
fitLoess <- loess(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked, data = summaryTableFirstDegAff, span = 0.2)
summary(fitLoess) #RSE is 0.00397 (seems good but I'm not sure)

loessVis = ggplot(data = summaryTableFirstDegAff, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitLoess)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "Comparing Carrier Risk for Masked Info and Full Family Information with Loess Regression", subtitle="First Degree Family Members Affected")

print(loessVis)
```

```{r loess model evaluation first deg aff}
# get predicted values for training data
summaryTableFirstDegAff$predicted <- predict(fitLoess)

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegAff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegAff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```


### First Degree Unaffected Families

```{r loess model fit and plot first deg unaff}
fitLoess <- loess(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked, data = summaryTableFirstDegUnaff, span = 0.2)
summary(fitLoess) 

loessVis = ggplot(data = summaryTableFirstDegUnaff, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitLoess)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "Comparing Carrier Risk for Masked Info and Full Family Information with Loess Regression", subtitle="First Degree Family Members Affected")

print(loessVis)
```

```{r loess model evaluation first deg unaff}
# get predicted values for training data

summaryTableFirstDegUnaff$predicted <- predict(fitLoess)

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegUnaff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegUnaff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```





## Median Polish

### All Families

```{r median polish model fit and plot}
fitMedianPolished <- rlm(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked, data = summaryTable)
print(summary(fitMedianPolished))

ggplot(data = summaryTable, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitMedianPolished)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "Comparing Carrier Risk for Masked Info and Full Family Information with Loess Regression")
```


```{r model evaluation median polish all families}
# get predicted values for training data

summaryTable$predicted <- predict(fitMedianPolished) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTable$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTable$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTable$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```

### First Degree Affected Families

```{r median polish model fit and plot first deg aff}
fitMedianPolished <- rlm(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked, data = summaryTableFirstDegAff)
print(summary(fitMedianPolished))

ggplot(data = summaryTableFirstDegAff, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitMedianPolished)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "Comparing Carrier Risk for Masked Info and Full Family Information with Loess Regression", subtitle="First Degree Family Members Affected")
```

```{r median polished model evaluation first deg aff}
# get predicted values for training data
summaryTableFirstDegAff$predicted <- predict(fitMedianPolished) #need to do the expit to undo the logit transformation

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegAff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegAff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegAff$probandMLH1Status) - summaryTableFirstDegAff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```

### First Degree Unaffected Families

```{r median polish model fit and plot first deg unaff}
fitMedianPolished <- rlm(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked, data = summaryTableFirstDegUnaff)
print(summary(fitMedianPolished))

ggplot(data = summaryTableFirstDegUnaff, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point(aes(color = probandMLH1Status)) +
  geom_line(aes(y = predict(fitMedianPolished)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "MLH1 Carrier Risk for Masked Info and Full Family Information with Median Polished Regression", subtitle="First Degree Family Members Unaffected")
```


```{r median polished model evaluation first deg unaff}
# get predicted values for training data

summaryTableFirstDegUnaff$predicted <- predict(fitMedianPolished) 

# calculate average predicted value
avg_adj <- mean(summaryTableFirstDegUnaff$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTableFirstDegUnaff$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTableFirstDegUnaff$probandMLH1Status) - summaryTableFirstDegUnaff$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```


# Ad Hoc Approach

```{r}
summaryTable$lowScoreIndicator <- ifelse(summaryTable$fullCarrierRisk < 0.1, 1, 0)
 

summaryTable$groupLowScore <- ifelse(summaryTable$lowScoreIndicator == 1, "0",
                                  ifelse(summaryTable$lowScoreIndicator == 1, "1","NA"))

fit <- lm(fullCarrierRisk ~ carrierRiskUnaffectedInfoMasked + lowScoreIndicator, data= summaryTable)
print(summary(fit))

ggplot(data = summaryTable, aes(x= carrierRiskUnaffectedInfoMasked, y= fullCarrierRisk)) +
  geom_point() +
  geom_line(aes(y = predict(fit)))+
  labs(x = "Carrier Risk Unaffected Info Masked", y= "Carrier Risk for Full Family Information", title = "MLH1 Carrier Risk for Masked Info and Full Family Information", subtitle="First Degree Family Members Unaffected")
# get predicted values for training data

summaryTable$predicted <- predict(fit)

# calculate average predicted value
avg_adj <- mean(summaryTable$predicted) #adjusted means adjusted via prediction model
avg_unadj <- mean(summaryTable$carrierRiskUnaffectedInfoMasked)
avg_true <- mean(as.numeric(summaryTable$probandMLH1Status))
# print average predicted value
print(str_interp("Average predicted (adjusted) carrier risk score: ${avg_adj}"))
print(str_interp("Average masked (unadjusted) carrier risk score: ${avg_unadj}"))
print(str_interp("Average number of true carriers of MLH1: ${avg_true}"))

mse_adj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$predicted)**2)
mse_unadj <- mean((as.numeric(summaryTable$probandMLH1Status) - summaryTable$carrierRiskUnaffectedInfoMasked)**2)
print(str_interp("Adjusted MSE: ${mse_adj}"))
print(str_interp("Unadjusted MSE: ${mse_unadj}"))
```

